{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-07-28 thesis/education.tex
small copyedits education

\begin{table}[h]
	\begin{center}
		% TODO: calc p values for this
		% note, smth like fisher's method which "combines independent studies into a meta p value"
		% cannot be applied for 2 reasons - 1, these are not independent since the same studence were
		% running landslide on diff test cases; 2, the hypotheses tested are different for each test
	\begin{tabular}{r||cc|cc|c}
		& \multicolumn{2}{c|}{Landslide users} & \multicolumn{2}{c|}{Non-users} & \\
		Bug name & \#correct & total & \#correct & total & $\Delta P(\mathsf{correct})$ \\
		\hline
		%Exit UAF  & 9	& 21	& 2	& 16	& $+30.4\%$ \\ % not filtering out did not run tej; also using "old" thing..?
		Exit UAF  & 11	& 21	& 0	& 16	& $+52.4\%$ \\
		% I spent more hours than I should have looking at this landslide-user-submitted-bug group.
		% Can't name the group here, but if you're me and have the analysis file,
		% it's the one from f17 with 3 mutex_test logs in their "u" snapshot.
		% It seems like a mysterious older version of mutex.c was present between their "a" version
		% (which gets fully verified very quickly btw) and their "u" version (for which landslide finds
		% the bug in just a few minutes, no way they could be running the same version; plus the reported
		% data races and so on in the id-log don't line up to the "u" mutex.c source code).
		% That version was correct, then they updated it (introducing the bug) and ran a bunch of other tests,
		% but didn't rerun mutex_test to check their update.
		Mutex     & 17	& 18	& 19	& 19	& $-05.6\%$ \\
		% nb. this is not filter out the did not run paras -- smae for all of the above/below
		% Paraguay  & 22	& 24	& 12	& 13	& $-0.64\%$ \\ % not filtering out did not run para
		Paraguay  & 20	& 21	& 14	& 16	& $+07.7\%$ \\
		% nb this is any downgrade related ding not just dgr tested ones
		% nb 1 is msising bc didn't impl rwlock at all
		%Downgrade & 17	& 24	& 8	& 12	& $+4.17\%$ \\ % not filtering out either did not run, or unrel ding
		Downgrade & 10	& 13	& 19	& 23	& $-05.7\%$ \\ % TODO verify these buggers by eye
		\hline
		% TODO: figure out if calculating a p value (or w/e) for this the same way is valid,
		% since they summands are arguably not independent? ..or maybe they are?
		Total     & 58	& 73	& 52	& 74	& $+09.2\%$ \\
	\end{tabular}
	\end{center}
	\caption{
		Correlation of student Landslide use with solving certain bugs in their final submission during Fall 2017 and Spring 2018 semesters.
		Students who used Landslide in general but whose usage snapshots showed they did not run the test case in question are counted as non-Landslide-users here.
		$\Delta P$ stands for the increased probability that a student submits a correct P2
		(with respect to each bug)
		given that they use Landslide, versus not using Landslide; i.e.,
		$P(\mathsf{correct}|\mathsf{landslide})-
		P(\mathsf{correct}|\mathsf{non\text{-}user})$.
		Note that the bottom row's totals double-count students,
		which only makes sense if you believe the incidences of each bug within a given submission are independent.
		}
	\label{tab:eval-common-bugs}
\end{table}

@@

Though I did not ask this question on the CMU survey,
my experience handling student questions suggests CMU students
also mostly wish for
better explanations of data races and
more detail and clarity in the preemption traces.
Though I present the formal definition of data races in the lecture
(\sect{\ref{sec:education-pebbles-recruiting}})
and refer back to it in Landslide's documentation and output,
showing a concrete example
in future iterations of the user guide
would go a long way to illustrate the abstract concepts.
The preemption traces could be improved by making it clearer that the stack trace in each cell of the table
represents executing the thread in question from wherever it previously left off (or its inception)
all the way until it reaches that stack trace, then preempting it to run another thread.
They could also easily report more diagnostic information;
for example, showing the sets of memory conflicts between each thread,
annotating the type of each preemption point (yield, mutex, or data race),
and/or indicating the adversarial memory access for each data race preemption point.

>>>>

\begin{table}[h]
	\begin{center}
		% TODO: calc p values for this
		% note, smth like fisher's method which "combines independent studies into a meta p value"
		% cannot be applied for 2 reasons - 1, these are not independent since the same studence were
		% running landslide on diff test cases; 2, the hypotheses tested are different for each test
	\begin{tabular}{r||cc|cc|c}
		& \multicolumn{2}{c|}{Landslide users} & \multicolumn{2}{c|}{Non-users} & \\
		Bug name & \#correct & total & \#correct & total & $\Delta P(\mathsf{correct})$ \\
		\hline
		%Exit UAF  & 9	& 21	& 2	& 16	& $+30.4\%$ \\ % not filtering out did not run tej; also using "old" thing..?
		Exit UAF  & 11	& 21	& 0	& 16	& $+52.4\%$ \\
		% I spent more hours than I should have looking at this landslide-user-submitted-bug group.
		% Can't name the group here, but if you're me and have the analysis file,
		% it's the one from f17 with 3 mutex_test logs in their "u" snapshot.
		% It seems like a mysterious older version of mutex.c was present between their "a" version
		% (which gets fully verified very quickly btw) and their "u" version (for which landslide finds
		% the bug in just a few minutes, no way they could be running the same version; plus the reported
		% data races and so on in the id-log don't line up to the "u" mutex.c source code).
		% That version was correct, then they updated it (introducing the bug) and ran a bunch of other tests,
		% but didn't rerun mutex_test to check their update.
		Mutex     & 17	& 18	& 19	& 19	& $-05.6\%$ \\
		% nb. this is not filter out the did not run paras -- smae for all of the above/below
		% Paraguay  & 22	& 24	& 12	& 13	& $-0.64\%$ \\ % not filtering out did not run para
		Paraguay  & 20	& 21	& 14	& 16	& $+07.7\%$ \\
		% nb this is any downgrade related ding not just dgr tested ones
		% nb 1 is msising bc didn't impl rwlock at all
		%Downgrade & 17	& 24	& 8	& 12	& $+4.17\%$ \\ % not filtering out either did not run, or unrel ding
		Downgrade & 10	& 13	& 19	& 23	& $-05.7\%$ \\ % TODO verify these buggers by eye
		\hline
		% TODO: figure out if calculating a p value (or w/e) for this the same way is valid,
		% since they summands are arguably not independent? ..or maybe they are?
		Total     & 58	& 73	& 52	& 74	& $+09.2\%$ \\
	\end{tabular}
	\end{center}
	\caption{
		Correlation of student Landslide use with solving certain bugs in their final submission during Fall 2017 and Spring 2018 semesters.
		Students who used Landslide in general but whose usage snapshots showed they did not run the test case in question are counted as non-Landslide-users here.
		$\Delta P$ stands for the increased probability that a student submits a correct P2
		(with respect to each bug)
		given that they use Landslide, versus not using Landslide; i.e.,
		$P(\mathsf{correct}|\mathsf{landslide})-
		P(\mathsf{correct}|\mathsf{non\text{-}user})$.
		Note that the totals in the bottom row double-count students,
		which only makes sense if you believe the incidences of each bug within a given submission are independent.
		}
	\label{tab:eval-common-bugs}
\end{table}

@@

Though I did not ask this question on the CMU survey,
my experience handling student questions in person
suggests CMU students
also mostly wish for
better explanations of data races and
more detail and clarity in the preemption traces.
Though I present the formal definition of data races in the lecture
(\sect{\ref{sec:education-pebbles-recruiting}})
and refer back to it in Landslide's documentation and output,
showing a concrete example
in future iterations of the user guide
would go a long way to illustrate the abstract concepts.
The preemption traces could be improved by making it clearer that the stack trace in each cell of the table
represents executing the thread in question from wherever it previously left off (or its inception)
all the way until it reaches that stack trace, then preempting it to run another thread.
They could also easily report more diagnostic information;
for example, showing the sets of memory conflicts between each thread,
annotating the type of each preemption point (yield, mutex, or data race),
and/or indicating the adversarial memory access for each data race preemption point.

### 2018-07-28 thesis/overview.tex
small copyedits education

In the Spring 2018 semester,
the Operating Systems class at Penn State University (henceforth \psuos and PSU, respectively)
offered the P2 thread library project as part of its curriculum.
Students in this class implemented P2
on a 6 week project timeline (compared to 2 weeks at CMU),
work alone rather than in pairs,
skip the {\tt swexn} automatic stack growth portion,
and rather than running their code with a reference Pebbles kernel binary in a simulator,
use the Pebwine emulation layer \cite{pebwine}
to run Pebbles-compatible program binaries in the Linux userspace.
Otherwise, the project is identical to CMU 15-410's P2.

>>>>

In the Spring 2018 semester,
the Operating Systems class at Penn State University (henceforth \psuos and PSU, respectively)
offered the P2 thread library project as part of its curriculum.
Students in this class implement P2
on a 6 week project timeline (compared to 2 weeks at CMU),
work alone rather than in pairs,
skip the {\tt swexn} automatic stack growth portion,
and rather than running their code with a reference Pebbles kernel binary in a simulator,
use the Pebwine emulation layer \cite{pebwine}
to run Pebbles-compatible program binaries in the Linux userspace.
Otherwise, the project is identical to CMU 15-410's P2.
